\sys: Automated Data-Parallel Shell Processing

### 2020-05-03 pldi/draft.tex

\heading{Discussion}
\kk{Hypothesis on results. Dish gives a great speedup if one step of
  the pipeline is what takes the most time. If the pipeline is long
  with many computation heavy tasks, task based parallelism gets you
  part of the way there, so we can't expect Dish to give performance
  benefits equal to the amount of parallelism that we give it.}

@@

\heading{Parallelization Command}
We compare \sys to \ttt{parallel} (v.20160422), a GNU utility for running other commands in parallel~\cite{Tange2011a}.
In the first experiment, we leverage \ttt{parallel}'s \ttt{--jobs} flag and redirect input from a file to 
Sequential execution takes \todo{XX}s, and \sys's parallel execution takes \todo{XX}s (speedup: \todo{XX}\%).
We note that this pipeline is harsh for \sys, in that most of the overhead comes from a single command, masking \sys's improvements in other parts of the pipeline.

>>>>

\heading{Discussion}
\kk{Below follows an important point about why scalability is more
  than great in the beginning and then gets worse.}
An interesting point is that \sys seems to achieve almost linear
speedup with small parallelism configurations. The reason why this
happens is not so straightforward though. \sys manages that because it
exploits task-based parallelism by having different processes to do
the \todo{map} and the \todo{reduce} phase. Due to task-based
parallelism in the generated dataflow graphs, \sys essentially ``uses
more parallelism'' than the given configuration. This can be seen in
all pipelines that are not completely stateless. For example, in the
\ttt{minimal\_sort} one-liner, \sys configured to have 8 parallelism
spawns 37 nodes (8 \ttt{tr} nodes, 8 \ttt{sort} nodes, 7 \ttt{sort -m}
nodes, and 14 \ttt{eager} nodes. This indicates that \sys achieves
optimal performance (without unnecessary process spawns and context
switches) in pipelines that contain pure commands by configuring it
with a smaller parallelism than the capabilities of the underlying
system -- in our case 16-32 for a 64 physical core system. This is
also the case if the input pipeline is very long and so there is
already task-based parallelism even in the sequential execution
(several of the Unix50 pipelines fall in this category).

\kk{Hypothesis on a scalability upper bound. Dish gives a great
  speedup if one step of the pipeline is what takes the most time. If
  the pipeline is long with many computation heavy tasks, task based
  parallelism gets you part of the way there, so we can't expect Dish
  to give performance benefits equal to the amount of parallelism that
  we give it.}

@@

\heading{\ttt{sort --parallel}}
We compare a single \todo{GNU} \ttt{sort} optimized by \sys with the
same \ttt{sort} with the \ttt{--parallel} flag set. The results are
shown in \Cref{fig:sort-parallel-comparison}. There are a couple
important points to note. First of all, \sys without the \ttt{eager}
optimization enabled performs comparably with \ttt{sort
  --parallel}. On the other hand, enabling the \ttt{eager}
optimization allows \sys to perform much better (almost \todo{2}
times) than \ttt{sort} with the \ttt{--parallel} flag. That is because
\sys with the \ttt{eager} optimization enabled essentially uses
intermediate buffers (on disk) between the merge phases. On the other
hand, \ttt{sort --parallel} sorts the input trying to minimize
additional memory (so doesn't allocate intermediate buffers). \kk{If
  we want, we can add here numbers where we allow sort --parallel to
  use additional ram for intermediate buffers (using the -S flag).} Of
course, it would be possible to manually create a longer pipeline and
use intermediate buffers to speed this up, but this defeats the
purpose of having a simple solution.

Finally, based on the comparison with \ttt{sort --parallel} it seems
that the scaleup limit that \sys reaches with sort is \todo{inherent
  and cannot be overpassed} \kk{We have to find a way to argue for the
  previous statement}. Because of that, all pipelines and scripts that
contain \ttt{sort} in our evaluation (several one-liners and unix50)
have this upper bound on scalability. \kk{Maybe move this paragraph in
  the one-liners section.}

\begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{\detokenize{./figs/sort_baseline_comparison_scaleup.pdf}}
  \caption{ \textbf{Comparison with \ttt{sort --parallel}.}
  }
  \vspace{-15pt}
  \label{fig:sort-parallel-comparison}
\end{figure}

\kk{There is a weird thing in the results. That the sequential one
  with --parallel=1 takes more than the \sys sequential one.}

\heading{GNU Parallel}
We compare \sys to \ttt{parallel} (v.20160422), a GNU utility for running other commands in parallel~\cite{Tange2011a}.
In the first experiment, we leverage \ttt{parallel}'s \ttt{--jobs} flag and redirect input from a file to 
Sequential execution takes \todo{XX}s, and \sys's parallel execution takes \todo{XX}s (speedup: \todo{XX}\%).
We note that this pipeline is harsh for \sys, in that most of the overhead comes from a single command, masking \sys's improvements in other parts of the pipeline.
